---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# atableExtra

[`atable`](https://CRAN.R-project.org/package=atable) is a package simplifying 
the task of creating tables of summary statistics 
similar to those found as the first table of a clinical trial publication. Although 
it is very flexible, utilising this flexibility involves writing new functions to 
calculate other measures and statistics. `atableExtra` provides some additional 
functions to simplify this task. `atableExtra` also changes some `atable` default 
settings.

## Installation

<!-- You can install the released version of atableExtra from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ```{r , eval=FALSE} -->
<!-- library(atableExtra) -->
<!-- ## basic example code -->
<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/) with: -->

```{r , eval=FALSE}
# install.packages("devtools")
devtools::install_github("aghaynes/atableExtra")
```

## Factors

`atableExtra` changes the default formatting of factors from `percentage (N)` to `N (percentage)`. 

```{r , eval=TRUE}
library(atable)
atable_options(format_to = "Console")
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
atable(mtcars, target_cols = "cyl")

library(atableExtra)
atable(mtcars, target_cols = "cyl")

```


## Numeric

`atableExtra` also provides a flexible approach to formatting numeric variables 
via the `numeric_stats` function and it's appropriate `format_statistics` method.
It's default settings mimic those from `atable`:

```{r}
# atable
atable(mtcars, target_cols = "mpg")
# atableExtra
atable_options(statistics.numeric = numeric_stats) # replace the default function
atable(mtcars, target_cols = "mpg")
```

It's easy to change the format though via the `numstats` argument that takes 
either numbers 1 to 3 (defaults to 1)...
```{r}
atable(mtcars, target_cols = "mpg", numstats = 1:3)
```

... or named characters...
```{r}
atable(mtcars, target_cols = "mpg", 
       numstats = c("Mean" = "{mean}", "Median" = "{q0.5}"))
```

Any quantile can be used (0th (= minimum), 25th, median, 75th and 100th (=maximum) are available by default) via the `quantiles` argument and referred to as e.g. `q0.1` for the 10th percentile.
```{r}
atable(mtcars, target_cols = "mpg", 
       quantiles = .1, 
       numstats = c("10th percentile" = "{q0.1}"))
```

The number of valid/missing observations can be turned off via the `missingformat` argument...

```{r}
atable(mtcars, target_cols = "mpg", missingformat = FALSE)
```
... or modified as per the `numstats` argument...
```{r}
atable(mtcars, target_cols = "mpg", missingformat = c("Missing" = "{Nmissing}"))
```

## Different statistics for objects of the same type

Sometimes it is preferable to have different summary statistics and tests for different variables of the same type. For example, normally distributed variables are often summarized by mean and standard deviation while non-normally distributed variables are typically summarized by quantiles.

To implement such a thing with `atable` can be long winded. You need to create a new class, provide methods for that class etc. `atableExtra` provides a set of functions specifically for this. Convert your variable to the numeric_np class with the `as.numeric_np` function and you can specify two sets of formats.

```{r}
mtcars$disp <- as.numeric_np(mtcars$disp)
class(mtcars$disp)
# atable(mtcars, target_cols = c("mpg", "disp"))
atable(mtcars, target_cols = c("mpg", "disp"), statistics.numeric_np = numeric_stats)
```






