---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# atableExtra

[`atable`](https://CRAN.R-project.org/package=atable) is a package simplifying 
the task of creating tables of summary statistics 
similar to those found as the first table of a clinical trial publication. Although 
it is very flexible, utilising this flexibility involves writing new functions to 
calculate other measures and statistics. `atableExtra` provides some additional 
functions to simplify this task. `atableExtra` also changes some `atable` default 
settings.

## Installation

<!-- You can install the released version of atableExtra from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ```{r , eval=FALSE} -->
<!-- library(atableExtra) -->
<!-- ## basic example code -->
<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/) with: -->

```{r , eval=FALSE}
# install.packages("devtools")
devtools::install_github("aghaynes/atableExtra")
```

## Factors

`atableExtra` changes the default formatting of factors from `percentage (N)` to `N (percentage)`. 

```{r , eval=TRUE}
library(atable)
atable_options(format_to = "Console")
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
atable(mtcars, target_cols = "cyl")

library(atableExtra)
atable(mtcars, target_cols = "cyl")

```


## Numeric

`atableExtra` also provides a flexible approach to formatting numeric variables 
via the `numeric_stats` function and it's appropriate `format_statistics` method.
It's default settings mimic those from `atable`:

```{r}
# atable
atable(mtcars, target_cols = "mpg")
# atableExtra
atable_options(statistics.numeric = numeric_stats) # replace the default function
atable(mtcars, target_cols = "mpg")
```

It's easy to change the format though via the `numstats` argument that takes 
either numbers 1 to 3 (defaults to 1)...
```{r}
atable(mtcars, target_cols = "mpg", numstats = 1:3)
```

... or named characters...
```{r}
atable(mtcars, target_cols = "mpg", 
       numstats = c("Mean" = "{mean}", "Median" = "{q0.5}"))
```

Any quantile can be used (0th (= minimum), 25th, median, 75th and 100th (=maximum) are available by default) via the `quantiles` argument and referred to as e.g. `q0.1` for the 10th percentile.
```{r}
atable(mtcars, target_cols = "mpg", 
       quantiles = .1, 
       numstats = c("10th percentile" = "{q0.1}"))
```

The number of valid/missing observations can be turned off via the `missingformat` argument...

```{r}
atable(mtcars, target_cols = "mpg", missingformat = FALSE)
```
... or modified as per the `numstats` argument...
```{r}
atable(mtcars, target_cols = "mpg", missingformat = c("Missing" = "{Nmissing}"))
```


